---
title: "Report Draft"
output: github_document
---



```{r, include= FALSE}
library(tidyverse)
library(gsheet)
data <- gsheet2tbl("https://docs.google.com/spreadsheets/d/1p-1ftljxC06sAmkSKfADT1E8P_0N4u5dlqQHmJGc-x8/edit?usp=sharing")

# write.csv(data, "raw_survey_results.csv")

names(data) <- c("time", "sex", "age", "satisfaction", "primary_language", "level_education", "STEM", "Years_off_school")
data$sex[data$sex=="NANANA"] <- NA
data[is.na(data)] <- "Others"
data <- data %>% 
  select(-time) %>% 
  mutate(Years_off_school = str_replace(Years_off_school, " years", ""),
         level_education = str_replace(level_education, "Masters.*", "Masters+"),
         primary_language = str_replace(primary_language, " languages", ""))
nms <- colnames(data)
data[nms] <- lapply(data[nms], as.factor)
data$satisfaction <- fct_relevel(data$satisfaction, c("very unhappy", "unhappy", "okay", "happy", "very happy"))
data$Years_off_school <- fct_relevel(data$Years_off_school, c("0-2", "3-5", "5-10", "10+"))

# write.csv(data, "cleaned_survey_results.csv")
```

```{r, include= FALSE}
data_sub <- data %>%
  mutate(satisfaction_level = ifelse(satisfaction == "very unhappy", 0,
                                     ifelse(satisfaction == "unhappy", 1,
                                            ifelse(satisfaction == "okay", 2,
                                                   ifelse(satisfaction == "happy", 3,4))))) %>%
  mutate(above_bachelor = ifelse(level_education == "Bachelors", 0 , 1))

data_sub$satisfaction_level <- as.double(data_sub$satisfaction_level)

Visualization <- function(data, conf, mode = "response"){
  if(mode == "response"){
    data_sub <- cbind(data[conf], data$satisfaction)
    names(data_sub) <- c("predictor", "satisfaction")
  
    p_pre_total <- data_sub %>% group_by(predictor) %>% summarise(total_count = n())
    p_pre_data <- data_sub %>% group_by(predictor, satisfaction) %>% summarise(count = n())
    p_pre_data <- left_join(p_pre_data, p_pre_total)
    p_pre_data <- p_pre_data %>%
     mutate(prop = count/total_count)
    
   count_plot <- p_pre_data %>%
      ggplot() +
      geom_bar(aes(x = predictor, y = count, fill = satisfaction), stat = "identity", position = 'dodge') +
      theme_bw() +
      labs(title = "Count plot", x = conf, y = "")
    
    prop_plot <- p_pre_data %>%
      ggplot() +
      geom_bar(aes(x = predictor, y = prop, fill = satisfaction), stat = "identity", position = 'dodge') +
      theme_bw() +
      labs(title = "Normalized proportion within group", x = conf, y = "")
    
   dist_pre <- p_pre_total %>%
     ggplot() +
     geom_bar(aes(x = predictor, y = total_count), stat = "identity") +
     theme_bw() +
     labs(title = "Count", x = conf, y = "")
    
  
  } else{
    data_sub <- cbind(data[conf], data$level_education)
    names(data_sub) <- c("conf", "level_education")
    
    p_pre_total <- data_sub %>% group_by(conf) %>% summarise(total_count = n())
    p_pre_data <- data_sub %>% group_by(conf, level_education) %>% summarise(count = n())
    p_pre_data <- left_join(p_pre_data, p_pre_total)
    p_pre_data <- p_pre_data %>%
     mutate(prop = count/total_count)
    
    count_plot <- p_pre_data %>%
      ggplot() +
      geom_bar(aes(x = conf, y = count, fill = level_education), stat = "identity", position = 'dodge') +
      theme_bw() +
      labs(title = "Count plot", x = conf, y = "")
      
    prop_plot <- p_pre_data %>%
      ggplot() +
      geom_bar(aes(x = conf, y = prop, fill = level_education), stat = "identity", position = 'dodge') +
      theme_bw() +
      labs(title = "Normalized proportion within group", x = conf, y = "")
    
   dist_pre <- p_pre_total %>%
     ggplot() +
     geom_bar(aes(x = conf, y = total_count), stat = "identity") +
     theme_bw() +
     labs(title = "Count", x = conf, y = "")
  
  }
  gridExtra::grid.arrange(count_plot, prop_plot,dist_pre, ncol = 2, nrow = 2)
}
```

```{r, include= FALSE}
glm_reg <- function(data, mode, conf, output = "model"){
  if (mode == "response"){
    data_sub <- cbind(data[conf], data$satisfaction_level)
    names(data_sub) <- c("confonder", "satisfaction_level")
    m <- glm(satisfaction_level ~ confonder, data=data_sub, family = "poisson")
    
  }else{
    data_sub <- cbind(data[conf], data$above_bachelor)
    names(data_sub) <- c("confonder", "above_bachelor")
    m <- glm(above_bachelor ~ confonder, data=data_sub, family = "poisson")
  }
  
  if(output == "model"){
    return(m)
  }else{
    return(summary(m))
  }
  # summary_m <- summary(m)
  
}
```

```{r, include = FALSE}
# model_test <- 
# model_test

# Visualization(data_sub, "age", "predictor")
```


## Overall Summary

#### The interaction between predictors and response

```{r}

# data_sub
model_overall <- glm(satisfaction_level ~ level_education, data = data_sub, family = 'poisson')
summary(model_overall)
```

#### The interaction between predictor, confonder and response 

```{r}

model_overall_conf <- glm(satisfaction_level ~ sex + age + level_education + STEM, data = data_sub, family = 'poisson')
summary(model_overall_conf)

```

#### Anova


## Confonders vs. Predictor

#### Age

We first explore the interaction between the age and our predictor of interest.

```{r}

Visualization(data_sub, "age", "predictor")
```


As we can observe from the EDA, different age group has distinct pattern for the proportion distribution of its own education level. For older people, they are intended to have master degree than yourger ones. However, we want to explore these pattern are significantly unique based on the hypothesis that **if age can significantly influence the predictor as a confunder**. Given that hypothesis, we set the age lower than 26 being the control group and there is no significant different between different age groups given the p-value. 

```{r}
m <- glm_reg(data = data_sub, mode = "predictor", conf = "age", output = "summary")
m
```


#### Sex

We discover similar results in the sex variable.

```{r}
Visualization(data_sub, "sex", "predictor")

```

```{r}

m <- glm_reg(data = data_sub, mode = "predictor", conf = "age", output = "summary")
m

```


#### STEM

```{r}
Visualization(data_sub, "STEM", "predictor")
```


```{r}

m <- glm_reg(data = data_sub, mode = "predictor", conf = "STEM", output = "summary")
m


```

## Confounders VS Response 


####  Age

```{r}

Visualization(data_sub, "age", "response")
```

##### H0: Age cannot significantly influence the response as a confunder 

```{r}
m_age <- glm_reg(data = data_sub, mode = "response", conf = "age", output = "summary")
m_age
```

Given p-value of all three groups of age are larger than 0.05, we cannot reject the null hypothesis. Therfore, age group 21-25 has no significant diffrent with other age groups and age cannot significantly influence the response as a confunder.  

####  Sex

```{r}

Visualization(data_sub, "sex", "response")
```

##### H0: Sex cannot significantly influence the response as a confunder 

```{r}
m_sex <- glm_reg(data = data_sub, mode = "response", conf = "sex", output = "summary")
m_sex
```
Given p-value of two groups of sex are larger than 0.05, we cannot reject the null hypothesis. Therfore, Female group has no significant different with other groups and sex cannot significantly influence the response as a confunder. 


####  STEM

```{r}

Visualization(data_sub, "STEM", "response")
```

##### H0: STEM cannot significantly influence the response as a confunder 

```{r}
m_stem <- glm_reg(data = data_sub, mode = "response", conf = "STEM", output = "summary")
m_stem
```

Given p-value of STEM are larger than 0.05, we cannot reject the null hypothesis. Therfore, STEMNo group has no significant different with the other group and STEM cannot significantly influence the response as a confunder. 









